---
title: "Setting Up the Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Setting Up the Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

```{r message = FALSE}
library(EMigD)
library(lubridate)
library(dplyr)
library(tibble)
```

## Setting Up the Data

The example shown below is one of many ways the data could be processed. The important components required is a `list` object containing home range kernels for each interval that is created for at least one individual.

For example, for a single individual if I wanted to break a one month period into 10-day intervals I would get a total of three 10-day intervals. These intervals should be in a `list` object, and each element within the `list` should represent a single interval.

The selected periods to be divided into intervals should be chosen so that they fall between migratory trips and don't overlap them. For example if individuals typically begins their migratory trip in spring and fall, the selected periods should not overlap spring and fall, and should be during the winter and summer seasons.

### Formatting the Data
Before moving into the data, we could first format the data so that we can go through the workflow smoothly using `EmigD_format`. This allows us to have a consistent column names as we move through the workflow. The arguments in `EmigD_format` consist of the necessary columns needed.  

```{r}
deer <- EmigD_format(deer, id_ = "ID", dt_ = "date", x_ = "utm_x", y = "utm_y",
                   year_ = "year", month_ = "month", jdate_ = "julian")
```


### Create Intervals

We created 10-day intervals for deer in the January and July. We first created a `start` column using the `floor_Date()` function to obtain the 10-day extents within the months. We included the 31st day (if a month had 31 days) in the third 10-day interval to prevent the creation of a 4th 10-day interval with only a single day in it.

### Intervals for January

```{r Jan Intervals}
jan <- deer %>%
  # Creates a start column assigning the first day in the 10-day interval in which
  # the date falls under (e.g., 01-03-2021 would be in the first 10-day interval
  # so the `floor_date` assigned to it would be 01-01-2021)
  mutate(start = floor_date(dt_, "10 days")) %>%
  # For any months that has 31 days, the 31st day would normally be assigned its 
  # own interval. The code below takes the 31st day and joins it with the 
  # previous interval. 
  mutate(start = if_else(day(start) == 31, start - days(10), start)) %>% 
  group_by(id_, start) %>% 
  filter(month_ == "1") %>% 
  group_split()
```

### Intervals for July

```{r July Intervals}
july <- deer %>%
  # Creates a start column assigning the first day in the 10-day interval in which
  # the date falls under (e.g., 01-03-2021 would be in the first 10-day interval
  # so the `floor_date` assigned to it would be 01-01-2021)
  mutate(start = floor_date(dt_, "10 days")) %>%
  # For any months that has 31 days, the 31st day would normally be assigned its 
  # own interval. The code below takes the 31st day and joins it with the 
  # previous interval. 
  mutate(start = if_else(day(start) == 31, start - days(10), start)) %>% 
  group_by(id_, start) %>% 
  filter(month_ == "7") %>% 
  group_split()

```

Elements in the list for one individual in January:

```{r}
head(jan[1:3])
```

We then attach names for each list element in `jan` and `july`. In this case, we include the `id_` and the `year_` for each element. This step is important and it plays an important role when using functions from the `EmigD` package.  

```{r}
names(jan) <- sapply(jan, function(x) paste(x$id_[1],
                                            x$year_[1], sep = '_'))

names(july) <- sapply(july, function(x) paste(x$id_[1],
                                              x$year_[1], sep = '_'))
```


## Create Tracks for Each Interval

Before we estimate LoCoH home range for each of these 10-day intervals, we first create objects of class `track_xyt` using `make_track()` from the `amt` package. This function creates a `track` using the `x_` and `y_` columns from `jan` and `july`. `lapply()` allows us to apply the specified function across a `list` object (e.g., `jan` and `july`)

```{r tracks, message = FALSE}
track_list_sum <- lapply(july, function(x) {
  amt::make_track(tbl = x, .x = x_, .y = y_, .t = dt_,
             uid = id_,
             # lat/long: 4326 (lat/long, WGS84 datum).
             # utm: crs = sp::CRS("+init=epsg:32612"))
             crs = 32612)
})
track_list_win <- lapply(jan, function(x) {
  amt::make_track(tbl = x, .x = x_, .y = y_, .t = dt_,
             uid = id_,
             crs = 32612)
})


```


## Creates LoCoHs Using the Tracks

Using the `track` created for the `jan` and `july`, we estimate type "a" LoCoH home range using `hr_locoh()` from the `amt` package. If you would like to learn more about the arguments of this function please refer to `?amt::hr_locoh` for more details.

```{r LoCoHs}
sum_locoh <- lapply(track_list_sum, function(tracks){
    # Calculate LoCoH a*
    dmat <- dist(tracks[, c("x_", "y_")])
    a <- max(dmat)

    # Fit LoCoH
    locoh <- try(amt::hr_locoh(
      x = tracks,
      levels = seq(0.1, 1, by = 0.1),
      keep.data = TRUE,
      n = a,
      type = "a",
      rand_buffer = 1e-05
    ))
    
    
})
win_locoh <- lapply(track_list_win, function(tracks){
    # Calculate LoCoH a*
    dmat <- dist(tracks[, c("x_", "y_")])
    a <- max(dmat)

    # Fit LoCoH
    locoh <- try(amt::hr_locoh(
      x = tracks,
      levels = seq(0.1, 1, by = 0.1),
      keep.data = TRUE,
      n = a,
      type = "a",
      rand_buffer = 1e-05
    ))
    
    
})
```

## Creates Isopleths from the LoCoHs

We then created isopleths from the LoCoH home range estimates using `hr_isopleths()`from the `amt` package.

```{r Isopleths}
sum_iso <- lapply(sum_locoh, function(x){
  amt::hr_isopleths(x)
})
win_iso <- lapply(win_locoh, function(x){
  amt::hr_isopleths(x)
})

```


## Rasterizes the Isopleths

After creating isopleths for each of the intervals, we have to rasterize them. Similar to the the previous functions, we use `lapply()` to apply our function across our `list` object. We first select the resolution for that we want for our rasters (`res`). We then create a template raster with a resolution of `res` (keep in mind that the speed the isopleths are `rasterize()` will be dependent on the specified resolution, with time of processing increasing with smaller resolutions). The projection of the template raster will match that of the specified isopleth.

After creating rasters for each interval, we then subtract all the isopleth values from 1, then normalize them.

```{r, results = "hide"}

res <- 250

sum_raster <- lapply(sum_iso, function(iso){
# Creates template raster
    template_raster <- raster::raster(iso,
      resolution = res, vals = 0,
      crs = sp::CRS(sf::st_crs(iso)[[2]])
    )
    # Rasterizes the isopleth
    raster_locoh <- raster::rasterize(iso,
      template_raster,
      field = "level",
      fun = "first"
    )

    # Now you want to subtract all the isopleth values from 1
    raster::values(raster_locoh) <- 1 - raster::values(raster_locoh)

    # Now normalize
    x <- raster_locoh / raster::cellStats(raster_locoh, sum)
})

win_raster <- lapply(win_iso, function(iso){
# Creates template raster
    template_raster <- raster::raster(iso,   
      resolution = res, vals = 0,
      crs = sp::CRS(sf::st_crs(iso)[[2]])
    )
    # Rasterizes the isopleth
    raster_locoh <- raster::rasterize(iso,
      template_raster,
      field = "level",
      fun = "first"
    )

    # Now you want to subtract all the isopleth values from 1
    raster::values(raster_locoh) <- 1 - raster::values(raster_locoh)

    # Now normalize
    x <- raster_locoh / raster::cellStats(raster_locoh, sum)
})

names(win_raster) <- sapply(jan, function(x) paste(x$id_[1],
                                            x$year_[1], sep = '_'))

names(sum_raster) <- sapply(july, function(x) paste(x$id_[1],
                                              x$year_[1], sep = '_'))

```

### Attach attributes to each list of rasters

We attach attributes to our list of rasters to allow for more flexibility in working with our lists of rasters.

```{r, results = "hide"}
# Add an attribute to hold the attributes of each list element
attributes(win_raster) <- data.frame(id = sapply(jan, function(x) paste(x$id_[1])),
                                      interval_start_date = sapply(jan, function(x) paste(x$start[1])),
                                      year = paste0(sapply(jan, function(x) paste(x$year_[1])))
)

# Check the attributes
attributes(win_raster)

# Add an attribute "tab" to hold the attributes of each list element
attributes(sum_raster) <- data.frame(id = sapply(july, function(x) paste(x$id_[1])),
                                      interval_start_date = sapply(july, function(x) paste(x$start[1])),
                                      year = paste0(sapply(july, function(x) paste(x$year_[1])))
)

# Check the attributes
attributes(sum_raster)

```

## Setting up input for Calculating EMDs in Environmental Space

The data set up above have specifically focused on setting up the data for calculating EMDs in geographical space. The code below demonstrates how you can set up the data to calculate the EMDs in environmental space.

### Creating 'RasterStacks' to calculate EMD and 'EM-Speeds in Environmental Space

#### Environmental Covariates without Temporal Variation

We first import the environmental covariate file (e.g., DEM layer), then we re-project the extent from latitude and longitude to UTMs to match those of the home range rasters.

```{r}
  #Load raster
  UtahDEM <- raster::raster(system.file("extdata", "Utah_DEM.tif", package = "EMigD"))
  out_proj <- "+proj=utm +zone=12 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
  UtahDEM@extent <- raster::extent(raster::projectExtent(UtahDEM, out_proj))
```

The `env_raster_stack()` function takes the environmental covariate raster and resamples it to the home range rasters. It then takes the resampled rasters and the home range rasters and constructs a object with class `RasterStack`. The steps following the creation of these `RasterStacks` is the steps under "Earth Mover's Distance", where the `RasterLayers` objects (i.e., `sum_raster` and `win_raster`) are replaced with the `RasterStack` objects (i.e., `sum_rst` and `win_rst`).

```{r}
sum_rst <- env_raster_stack(UtahDEM, sum_raster)
win_rst <- env_raster_stack(UtahDEM, win_raster)

```

#### Environmental Covariates with Temporal Variation

For environmental covariates that vary across time (e.g., temperature, snow depth, foraging resources, precipitation, etc.), we first assign the file names of which the covariates are housed to `list_names`. In our example, we use Rangeland Analysis Platform (RAP) data in 2013 and 2014 that can be an indicator of forage resources on the landscape.

```{r}

list_names <- list.files(path=system.file("extdata", package = "EMigD"),
                 pattern=c("RAP","*.tif"), all.files= FALSE,
                 full.names = TRUE) 

basename(list_names)

```

We then read in those files into R as `RasterLayer` objects. For each of those `list` elements in `env_lists` we assign the path names for each of those files. Using the assigned names, we created `attributes` for our list with a `year` column. The `year` column may change depending on the temporal scale of your environmental covariate. For example, if we had environmental covariates that changed daily, we can first aggregate the daily covariates based on the intervals. Instead of using `year`, we would then use the start date of the intervals.

```{r}
# Rasterize the environmental covariates
env_list <- lapply(list_names, raster::raster)

names(env_list) <- sapply(list_names, function(x) paste(x))

# Uses the name of the environemntal covariate file to assign attributes 
attributes(env_list) <- data.frame(year = trimws(basename(names(env_list)), 
                                                 whitespace = "_.*"))

attributes(env_list)

```

We use `env_raster_stack` with the addition of the `attribute` argument to stack environmental covariate layers with our home range rasters. The input for the `attribute` argument should be a attribute component that is the same in both of the `list` objects (i.e., env_list and sum_raster/win_raster) with the same unique values. For example, both `env_list` and `sum_raster`/`win_raster` have a attribute component labelled `year` and both have the same unique values (e.g., 2013 and 2014).

NOTE: Currently this feature assumes that only environmental layers relevant to `sum_raster` or `win_raster` are being used and are in the same folder(i.e., excluding any `year` that are not present in `sum_raster` and/or `win_raster`.)

```{r}
sum_rst_temporal <- env_raster_stack(env_list, sum_raster, attribute = "year")
win_rst_temporal <- env_raster_stack(env_list, win_raster, attribute = "year")
```
