---
title: "Different intervals between seasons"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Different intervals between seasons}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

```{r message = FALSE}
library(EMigD)
library(lubridate)
library(dplyr)
library(tibble)
```

## Setting Up the Data

### Formatting the Data
Before moving into the data, we could first format the data so that we can go through the workflow smoothly using `EmigD_format`. This allows us to have a consistent column names as we move through the workflow. The arguments in `EmigD_format` consist of the necessary columns needed.  

```{r}
deer <- EmigD_format(deer, id_ = "ID", dt_ = "date", x_ = "utm_x", y = "utm_y",
                   year_ = "year", month_ = "month", jdate_ = "julian")
```


### Create Intervals

We created 10-day intervals for deer in the January and July. We first created a `start` column using the `floor_Date()` function to obtain the 10-day extents within the months. We included the 31st day (if a month had 31 days) in the third 10-day interval to prevent the creation of a 4th 10-day interval with only a single day in it.

### Intervals for January

```{r Jan Intervals}
jan <- deer %>%
  # Creates a start column assigning the first day in the 10-day interval in which
  # the date falls under (e.g., 01-03-2021 would be in the first 10-day interval
  # so the `floor_date` assigned to it would be 01-01-2021)
  mutate(start = floor_date(dt_, "10 days")) %>%
  # For any months that has 31 days, the 31st day would normally be assigned its 
  # own interval. The code below takes the 31st day and joins it with the 
  # previous interval. 
  mutate(start = if_else(day(start) == 31, start - days(10), start)) %>% 
  group_by(id_, start) %>%
  filter(month_ == "1") %>% 
  group_split()
```

Let's remove some list elements from the `jan` interval so that we have 
different number of intervals between `jan` and `july`.

```{r}
jan <- jan[-c(17,18)]
```

### Intervals for July

```{r July Intervals}
july <- deer %>%
  # Creates a start column assigning the first day in the 10-day interval in which
  # the date falls under (e.g., 01-03-2021 would be in the first 10-day interval
  # so the `floor_date` assigned to it would be 01-01-2021)
  mutate(start = floor_date(dt_, "10 days")) %>%
  # For any months that has 31 days, the 31st day would normally be assigned its 
  # own interval. The code below takes the 31st day and joins it with the 
  # previous interval. 
  mutate(start = if_else(day(start) == 31, start - days(10), start)) %>% 
  group_by(id_, start) %>%
  filter(month_ == "7") %>% 
  group_split()

```

### Matching Intervals

Before we can begin matching the intervals, we first assign attributes to the two list (`jan` and `july`). We name this attribute `match`, and within this attribute table, we have two columns (`id` and `interval_`). We can use `check` to conveniently look our attribute table. 
```{r}
# Add an attribute to hold the attributes of each list element
attr(jan, "match") <- data.frame(id = sapply(jan, function(x) paste(x$id_[1])),
                                  interval_ = sapply(jan, function(x) paste(x$start[1]))
)

# Check the attributes
check(jan, "match")

# Add an attribute "tab" to hold the attributes of each list element
attr(july, "match") <- data.frame(id = sapply(july, function(x) paste(x$id_[1])),
                                  interval_ = sapply(july, function(x) paste(x$start[1]))
) 

# Check the attributes
check(july, "match")
```


Using the `match_intervals` function we can match the number of intervals in 
`july` to those in `jan`. The `attribute_name` is the name of the attribute that was added. We then include the two list that we want to match. The `id_col` and the `interval_col` are the two columns created for the added attribute. 

```{r Matching Intervals}

match_list <- match_intervals(attribute_name = "match", list1 = july, list2 = jan, id_col = "id", 
                       interval_col = "interval_")

```

The output for `match_ouput` will be a list that contains two new lists. These two new list are the matching lists for the two specified in the arguments. These can be re-assigne to allow for readability or it can be used as is. 

```{r}
july <- match_list[[1]]
jan <- match_list[[2]]
```



