---
title: "Different intervals between seasons"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Different intervals between seasons}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

```{r message = FALSE}
library(EMigD)
library(lubridate)
library(dplyr)
library(tibble)
```

## Setting Up the Data

### Create Intervals

We created 10-day intervals for deer in the January and July. We first created a `start` column using the `floor_Date()` function to obtain the 10-day extents within the months. We included the 31st day (if a month had 31 days) in the third 10-day interval to prevent the creation of a 4th 10-day interval with only a single day in it.

### Intervals for January

```{r Jan Intervals}
jan <- deer %>%
  # Creates a start column assigning the first day in the 10-day interval in which
  # the date falls under (e.g., 01-03-2021 would be in the first 10-day interval
  # so the `floor_date` assigned to it would be 01-01-2021)
  mutate(start = floor_date(date, "10 days")) %>%
  # For any months that has 31 days, the 31st day would normally be assigned its 
  # own interval. The code below takes the 31st day and joins it with the 
  # previous interval. 
  group_by(ID) %>% 
  mutate(start = if_else(day(start) == 31, start - days(10), start)) %>% 
  group_by(start, .add = TRUE) %>%
  filter(month == "1") %>% 
  group_split()
```
Let's remove some list elements from the `jan` interval so that we have 
different number of intervals between `jan` and `july`. 
```{r}
jan <- jan[-c(17,18)]
```

### Intervals for July

```{r July Intervals}
july <- deer %>%
  # Creates a start column assigning the first day in the 10-day interval in which
  # the date falls under (e.g., 01-03-2021 would be in the first 10-day interval
  # so the `floor_date` assigned to it would be 01-01-2021)
  mutate(start = floor_date(date, "10 days")) %>%
  # For any months that has 31 days, the 31st day would normally be assigned its 
  # own interval. The code below takes the 31st day and joins it with the 
  # previous interval. 
  group_by(ID) %>% 
  mutate(start = if_else(day(start) == 31, start - days(10), start)) %>% 
  group_by(start, .add = TRUE) %>%
  filter(month == "7") %>% 
  group_split()

```

### Assign names to the list elements
The names that are assigned to each list element is specific to that interval.
In this case, we use the `ID` and `start` columns in the list element to assign 
a name. These names are important because they are later used to determine which 
elements are missing between the two list objects.

```{r Names for intervals}
names(jan) <- sapply(jan, function(x) paste(x$ID[1],
                                            x$start[1], sep = '_'))

names(july) <- sapply(july, function(x) paste(x$ID[1],
                                              x$start[1],sep = '_'))
                                              
```

### Matching Intervals
Using the `match_intervals` function we can match the number of intervals in 
`july` to those in `jan`. In this case, the first argument `list1` is the 
reference list for the second argument `list2`. 
```{r Matching Intervals}
july <- match_intervals(list1 = jan, list2 = july)

```
